decompr: GVC decomposition in R

### 2015-02-02 draft.tex
recompile

\section{Introduction}
\epigraph{Hello....is is me you're looking for?}{Lionel Richie}
Global Value Chains (GVCs) have become a central topic in research on trade and development policy.
Both policy makers and academia increasingly value the growth opportunities GVCs offer to global trade and, especially, to developing countries.
However, analysing this phenomenon empirically requires complex matrix manipulations, since the relevant data is only available in the form of gross flows.
This package enables researchers with little background in matrix algebra and linear programming to easily derive standard GVC indicators for statistical analysis.

The \textbf{decompr} package uses Inter-Country Input-Output tables (ICIOs), such as those published by the OECD and WTO (TiVA), % CITE!! 
the World Input Output Database \parencite{timmer2012world},  or national statistics bureaus, as input.
These tables state supply and demand relationships in gross terms between industries within and across countries.

For instance, let us look at the example of the leather used in German manufactured car seats.
The ICIOs quantify the value of inputs that the Turkish leather and textiles industry
supplies to the German transport equipment industry.

The problem of these gross trade flows, is that they do not reveal how much of the value was added in the supplying industry,
and how much of the value was added in previous stages of production, performed by other industries or even countries.

The \textbf{Leontief} decomposition of gross trade flows solves this problem by reallocating the value of intermediate goods used by industries to the original producers.
In our example, the use of Argentinian agricultural produce (raw hides) is subtracted from the Turkish leather industry and added to the Argentinian agricultural industry.

The \textbf{Wang-Wei-Zhu} (henceforth WWZ) decomposition goes a step further by not only revealing the source of the value added, but also breaking down exports into different categories, according to final usage and destination.
It implements the theoretical work of \textcite{wang2014quantifying}.
The high-level divisions in this framework are listed below.

@@

>>>>

\section{Introduction}
\epigraph{Hello?\\...\\ is it me you're looking for?}{Lionel Richie}
Global Value Chains (GVCs) refer to the quickly expanding internationalization of production networks.
Most goods we use nowadays consist of parts that are sourced from different corners of the planet 
and are assembled across different continents. A popular example of this development is the Iphone,
which uses inputs from at least five countries (USA, China, Germany, Taiwan, South Korea) and is assembled in two (USA and China).
This has made GVCs a central topic in research on trade and development policy.
Both policy makers and academia increasingly value the growth opportunities GVCs offer to global trade and, 
especially, to developing countries.
However, analysing this phenomenon empirically requires complex matrix manipulations, 
since the relevant data is only available in the form of gross flows.
This package enables researchers with little background in matrix algebra and linear programming to easily derive standard GVC indicators for statistical analysis.

The \textbf{decompr} package uses Inter-Country Input-Output tables (ICIOs), such as those published by the OECD and WTO (TiVA), % CITE!! 
the World Input Output Database \parencite{timmer2012world},  or national statistics bureaus, as input.
These tables state supply and demand relationships in gross terms between industries within and across countries.
For instance, let us look at the example of the leather used in German manufactured car seats.
The ICIOs quantify the value of inputs that the Turkish leather and textiles industry
supplies to the German transport equipment industry.

The problem of these tables measuring gross trade flows, 
is that they do not reveal how much of the value was added in the supplying industry,
and how much of the value was added in previous stages of production, performed by other industries or even countries.
The \textbf{Leontief} decomposition of gross trade flows solves this problem by reallocating the value of intermediate goods used by industries to the original producers.
In our example, the use of Argentinian agricultural produce (raw hides) is subtracted from the Turkish leather industry and added to the Argentinian agricultural industry.

The \textbf{Wang-Wei-Zhu} (henceforth WWZ) decomposition goes a step further by not only revealing the source of the value added,
but also breaking down exports into different categories, according to final usage and destination.
It implements the theoretical work of \textcite{wang2014quantifying}.
The main categories in this framework are listed below.

@@

GVCs describe the increasingly international organization of production structures.
As more and more regional trade agreements come into force, 
which drive down trade costs and harmonize product standards,
it becomes more and more attractive for firms to outsource certain tasks of their production lines.
Research on international trade analysing this development evolves quickly and reveals important implications of GVCs for economic growth and competitiveness.  
\textbf{decompr} aims at facilitating this reasearch by simplifying the calculation of standard GVC indicators.
The purpose is to accelerate the research and, especially, to make it accesible to a wider audience.

% Baast, enter part on how the package is flexible and can be adapted

Since GVCs constitute a fairly new field of research, there are many ways forward for its analysis. 
The next central step is to examine both in theory and empirically how GVC participation affects real economic activity. 
More specifically, it is very relevant to look at how, for instance, 
employment and economic growth react when countries join GVCs and what the factors are that determine a successful relationship. 
From the standpoint of developing and emerging countries a very interesting question is if GVCs simplify industrialization 
and the formation of comparative advantage 
while high-income countries might look for an additional push for their stagnating post-crisis economies. 
We hope that \textbf{decompr} can play a part in this field and promote it.
